Locality and Computational Reliability of Linear Response Calculations for Molecular Systems

### 2019-02-11 AnalyticStructure_main.tex
fig3 converted in eps and ref to supplementary material

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!t]
\begin{tikzpicture}[scale=0.78]
% axes
\draw[->] (0,0) -- (8,0);
\draw[->] (0,0) -- (0,6);
% orizontal lines
\draw[dashed,thick,color=gray] (0,5) -- (2,5);
\draw[very thick](2,5) -- (8,5);
\draw[dashed,thick,color=gray] (0,3) -- (4,3);
\draw[very thick] (4,3) -- (8,3);
\draw[dashed,thick,color=gray] (0,1) -- (6,1);
\draw[very thick] (6,1) -- (8,1);
% % vertical lines
\draw[dashed,very thin,color=black] (2,0) -- (2,5);
\draw[dashed,very thin,color=black] (4,0) -- (4,3);
\draw[dashed,very thin,color=black] (6,0) -- (6,1);
% % threshold levels labels
\node[left] at (0,5) {$p=h$};
\node[left] at (0,3) {$p$};
\node[left] at (0,1) {$p=1$};
\node[below] at (2,0) {$|\eps_h|$};
\node[below] at (4,0) {$|\eps_p|$};
\node[below] at (6,0) {$|\eps_1|$};
\node[below] at (4,-0.5) {excitation energy};
% bars at thresholds
%\foreach \Point in {(2,5),(4,3),(6,1)}{
%   \node at \Point {{\textbar}};%$\times$};
%}
\draw[very thick] (2,4.9) -- (2,5.1);
\draw[very thick] (4,2.9) -- (4,3.1);
\draw[very thick] (6,0.9) -- (6,1.1);
% dotted vertical lines
%\draw[dotted,thick,color=black] (-0.2,3.6) -- (-0.2,4.4);
%\draw[dotted,thick,color=black] (-0.2,1.6) -- (-0.2,2.4);
% discrete excitations
\foreach \Point in {(0.5,5),(1.0,5),(1.8,5),(0.8,3),(1.3,3),(1.8,3),(2.4,3),(2.8,3),(3.4,3),(3.8,3),(0.7,1),(1.5,1),(1.9,1),(2.6,1),(2.9,1),(3.5,1),(4.1,1),(4.7,1),(5.7,1)}{
    \node at \Point {\textbullet};
}
\end{tikzpicture}
\caption{\label{ExcitationLandscape} Excitation landscape. Excitations are split in horizontal lines according to the threshold levels $p$. Filled bullets refer to localized
excitations whereas the thick black lines describe delocalized ones belonging a continuum spectrum.}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The susceptibility functional has therefore a peculiar analytic structure made of multiple threshold energies, one for each of the ionization potentials associated with the occupied states. If an excitation $\op E_a$ has an energy $\Omega_a < |\eps_a|$ with $\eps_a = \mathrm{max}\left(\eps_p\right)_{p\in a}$, it may belong to a discrete spectrum; otherwise it \emph{has} to belong to a continuum of states.
The typical structure of the excitation landscape is depicted in figure \ref{ExcitationLandscape}, once again assuming that the perturbed potential $\op V'[\op E]$ gives rise to a bound state when applied to $\ket{\psi_p}$. Below the first ionization energy $|\eps_h|$ only localized excitations are present. Conversely, in the intermediate energy interval that extends from $|\eps_h|$ to the absolute value of the deepest occupied orbital, other discrete excitations may exist, and are embedded in a continuum of delocalized excitations. For energies higher than $|\eps_1|$ \emph{no localized} excitations are anymore possible,
regardless of the particular behavior of $\op V'$: the excitation landscape in this region contains a continuum set of delocalized states.

@@

We have analyzed the fulfillment of Eq.~\eqref{sminus2} for molecules of various
sizes and symmetries.
In Fig.~\ref{co_AlphaExc} we show explicitly such a comparison for the $CO$ and $C_{60}$ molecules.
We notice that, in both the cases, a considerable number of pseudo-continuum excitations is required
to find the correct result; the sector of the discrete excitations appears to be largely insufficient to express the 
reference fluctuation state. 
%In particular, in the case of $CO$, the contribution of localized excitations is even negligible w.r.t. the one coming from the continuum sector.
In particular, in the case of $CO$, the complete excitation set is able to reproduce the reference value of $\alpha_{zz}$, but the contribution of localized sector is negligible w.r.t. the one coming from the continuum one. Instead, for the $C_{60}$, the localized sector provides a more significant contribution but a higher number of the psuedo-continuum excitations should be included in the sum of \eqref{sminus2} to satisfy the $S_{-2}$ sum-rule.

@@

%This scenario can be also found for the other molecules presented in Table \ref{alphaTable};
This scenario is confirmed by the results for the other molecules presented in Table \ref{alphaTable};
varying the molecule size and symmetry does not modify the qualitative behavior of the excitation landscape (for further details
about the dependence of the excitation landscape on the size of the molecule see the supplementary material). In addition the
localized sector of the excitations alone is largely insufficient to express $\alpha_{ij}$.
Such considerations are independent of the number of atoms of the molecule, as the range in $\omega$ needed to
achieve satisfactory agreement is largely above the last valence ionization threshold, which is
an intensive quantity.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure}[!t]
% \begin{tikzpicture}[scale=0.78]
% % axes
% \draw[->] (0,0) -- (8,0);
% \draw[->] (0,0) -- (0,6);
% % orizontal lines
% \draw[dashed,thick,color=gray] (0,5) -- (2,5);
% \draw[very thick](2,5) -- (8,5);
% \draw[dashed,thick,color=gray] (0,3) -- (4,3);
% \draw[very thick] (4,3) -- (8,3);
% \draw[dashed,thick,color=gray] (0,1) -- (6,1);
% \draw[very thick] (6,1) -- (8,1);
% % % vertical lines
% \draw[dashed,very thin,color=black] (2,0) -- (2,5);
% \draw[dashed,very thin,color=black] (4,0) -- (4,3);
% \draw[dashed,very thin,color=black] (6,0) -- (6,1);
% % % threshold levels labels
% \node[left] at (0,5) {$p=h$};
% \node[left] at (0,3) {$p$};
% \node[left] at (0,1) {$p=1$};
% \node[below] at (2,0) {$|\eps_h|$};
% \node[below] at (4,0) {$|\eps_p|$};
% \node[below] at (6,0) {$|\eps_1|$};
% \node[below] at (4,-0.5) {excitation energy};
% % bars at thresholds
% %\foreach \Point in {(2,5),(4,3),(6,1)}{
% %   \node at \Point {{\textbar}};%$\times$};
% %}
% \draw[very thick] (2,4.9) -- (2,5.1);
% \draw[very thick] (4,2.9) -- (4,3.1);
% \draw[very thick] (6,0.9) -- (6,1.1);
% % dotted vertical lines
% %\draw[dotted,thick,color=black] (-0.2,3.6) -- (-0.2,4.4);
% %\draw[dotted,thick,color=black] (-0.2,1.6) -- (-0.2,2.4);
% % discrete excitations
% \foreach \Point in {(0.5,5),(1.0,5),(1.8,5),(0.8,3),(1.3,3),(1.8,3),(2.4,3),(2.8,3),(3.4,3),(3.8,3),(0.7,1),(1.5,1),(1.9,1),(2.6,1),(2.9,1),(3.5,1),(4.1,1),(4.7,1),(5.7,1)}{
%     \node at \Point {\textbullet};
% }
% \end{tikzpicture}
% \caption{\label{ExcitationLandscape} Excitation landscape. Excitations are split in horizontal lines according to the threshold levels $p$. Filled bullets refer to localized
% excitations whereas the thick black lines describe delocalized ones belonging a continuum spectrum.}
% \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!t]
\includegraphics[scale=0.12]{Fig3_ExcitationLandscape.eps}
\caption{\label{ExcitationLandscape} Excitation landscape. Excitations are split in horizontal lines according to the threshold
levels $p$. Filled bullets refer to localized excitations whereas the thick black lines describe delocalized ones belonging a
continuum spectrum.}
\end{figure}

The susceptibility functional has therefore a peculiar analytic structure made of multiple threshold energies, one for each of the ionization potentials associated with the occupied states. If an excitation $\op E_a$ has an energy $\Omega_a < |\eps_a|$ with $\eps_a = \mathrm{max}\left(\eps_p\right)_{p\in a}$, it may belong to a discrete spectrum; otherwise it \emph{has} to belong to a continuum of states.
The typical structure of the excitation landscape is depicted in Fig.~\ref{ExcitationLandscape}, once again assuming that the perturbed potential $\op V'[\op E]$ gives rise to a bound state when applied to $\ket{\psi_p}$. Below the first ionization energy $|\eps_h|$ only localized excitations are present. Conversely, in the intermediate energy interval that extends from $|\eps_h|$ to the absolute value of the deepest occupied orbital, other discrete excitations may exist, and are embedded in a continuum of delocalized excitations. For energies higher than $|\eps_1|$ \emph{no localized} excitations are anymore possible,
regardless of the particular behavior of $\op V'$: the excitation landscape in this region contains a continuum set of delocalized states.

@@

We have analyzed the fulfillment of Eq.~\eqref{sminus2} for molecules of various
sizes and symmetries.
In Fig.~\ref{co_AlphaExc} we show explicitly such a comparison for the $CO$ and $C_{60}$ molecules.
We notice that, in both the cases, a considerable number of pseudo-continuum excitations is required
to find the correct result; the sector of the discrete excitations appears to be largely insufficient to express the 
reference fluctuation state. 
In particular, in the case of $CO$, the complete excitation set is able to reproduce the reference value of $\alpha_{zz}$, but the contribution of localized sector is negligible w.r.t. the one coming from the continuum one. Instead, for the $C_{60}$, the localized sector provides a more significant contribution but a higher number of the psuedo-continuum excitations should be included in the sum of \eqref{sminus2} to satisfy the $S_{-2}$ sum-rule.

@@

This scenario is confirmed by the results for the other molecules presented in Table \ref{alphaTable}; varying the molecule size
and symmetry does not modify the qualitative behavior of the excitation landscape (see Supplementary Material at~\footnote{\texttt{URL insert by the publisher}} for further details about the dependence of the excitation landscape on the size of the molecule). 
In addition the localized sector of the excitations alone is largely insufficient to express $\alpha_{ij}$. Such considerations 
are independent of the number of atoms of the molecule, as the range in $\omega$ needed to achieve satisfactory agreement is largely above the last valence ionization threshold, which is an intensive quantity.
